Relative parabolic factorizations of Chevalley groups

### 2015-04-28 bass-kolster.tex
BK small fixes

\begin{thm}\label{thm:BassKolster}
Let $\Phi$ be a classical root system of rank $\geqslant2$, let $R$ be arbitary commutative ring and $I$ be an ideal, satisfying the following condition:
\begin{itemize}
\item $\Phi=\rA_\ell$: $\sr(I)\leqslant\ell$,
\item $\Phi=\rB_\ell$: $\asr(R,I)\leqslant\ell-1$,
\item $\Phi=\rC_\ell$: $\sr(I)\leqslant2\ell$,
\item $\Phi=\rD_\ell$: $\asr(R,I)\leqslant\ell-1$.
\end{itemize}
Then the principal congruence subgroup $\G(\Phi,R,I)$ admits the following relative version of Bass---Kolster decomposition:
\[\G(\Phi,R,I)=\U(\Sigma^+,I)\ \U(\Sigma^-,I)\cdot \Z\cdot \U(\Sigma^-,I)\ \U(\Sigma^+,I)\cdot \G(\Delta,R,I), \]
where $\Sigma^\pm=\Phi^\pm\setminus\Delta$ and $\Z=\{ z_{-\alpha_\mathrm{max}}(r,1)\mid r\in I \}$.
\end{thm}
\begin{proof}
Let $g$ be an element of $\G(\Phi, R, I)$ and $e_1$ be the highest weight vector of the standard vector representation of $\G(\Phi, R, I)$.
It is clear that vector $v=g\cdot e_1$ is $I$-unimodular.
%TODO: Mention Chevalley---Matsumoto lemma in the introduction
In view of Chevalley---Matsumoto lemma it suffices to show that $\Z \cdot \U(\Phi^-, I) \cdot \U(\Phi^+, I) \cdot g$ contains an element $g$ whose matrix entry at position $(1,1)$ is equal to $1$.
We will need to analyze several cases.

>>>>

\begin{thm}\label{thm:BassKolster}
Let $\Phi$ be a classical root system of rank $\geqslant2$, let $R$ be arbitary commutative ring and $I$ be an ideal, satisfying the following assumption:
\begin{itemize}
\item $\Phi=\rA_\ell$: $\sr(I)\leqslant\ell$,
\item $\Phi=\rB_\ell$: $\asr(R,I)\leqslant\ell-1$,
\item $\Phi=\rC_\ell$: $\sr(I)\leqslant2\ell-1$,
\item $\Phi=\rD_\ell$: $\asr(R,I)\leqslant\ell-1$.
\end{itemize}
Then the principal congruence subgroup $\G(\Phi,R,I)$ admits the following relative version of Bass---Kolster decomposition:
\[\G(\Phi,R,I)=\U(\Sigma^+,I)\ \U(\Sigma^-,I)\cdot \Z\cdot \U(\Sigma^-,I)\ \U(\Sigma^+,I)\cdot \G(\Delta,R,I), \]
where $\Sigma^\pm=\Phi^\pm\setminus\Delta$ and $\Z=\left\{ z_{-\alpha_\mathrm{max}}(r,1)\mid r\in I \right\}$.
\end{thm}
\begin{proof}
Let $g$ be an element of $\G(\Phi, R, I)$ and $e_1$ be the highest weight vector of the standard vector representation of $\G(\Phi, R, I)$.
It is clear that vector $v=g\cdot e_1$ is $I$-unimodular.
%TODO: Mention Chevalley---Matsumoto lemma in the introduction
In view of Chevalley---Matsumoto lemma it suffices to show that $\Z \cdot \U(\Phi^-, I) \cdot \U(\Phi^+, I) \cdot g$ contains an element $g$ whose matrix entry at position $(1,1)$ is equal to $1$.
We will need to analyze several cases.
